---
title: "project"
author: "Ankita Mishra"
date: "2024-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# libraries import
library(ggplot2)
library(GGally)
library(corrplot)
library(readr)
library(leaps)
library(car)
library(MASS)
library(psych)
library(dplyr)
library(glmnet)
library(pracma)
```
```{r}
coffee = read.csv('Coffee_Chain_Sales .csv')
head(coffee)
names(coffee)[sapply(coffee, is.integer)]
names(coffee)[sapply(coffee, is.character)]
```
```{r}
# check for missing data
sum(is.na(coffee))
```
```{r}
summary(coffee)
```
```{r}
unique(coffee$Market_size)
unique(coffee$Market)
unique(coffee$Product_line)
unique(coffee$Product_type)
unique(coffee$Product)
unique(coffee$State)
unique(coffee$Type)
```


# EDA and log transformations done by [Manoj](https://github.com/MK-Github03)
# Coffee Chain Sales Data Analysis - Exploratory Data Analysis (EDA)
# This script provides an initial exploration of the Coffee Chain Sales dataset, 
# focusing on summary statistics, missing data, correlations, outlier detection, 
# and the distributions of numeric variables. These steps help understand the data's 
# characteristics, identify potential issues, and guide subsequent analysis and 
# modeling efforts.

install.packages("ggplot2")
install.packages("dplyr")
install.packages("corrplot")

library(ggplot2)
library(dplyr)
library(corrplot)

# Loading the dataset
dataset <- read.csv('Coffee_Chain_Sales.csv')

# Summary 
summary(dataset)

# Check for missing values
missing_data <- colSums(is.na(dataset))
print(missing_data)

#missing values found : 0, fixed by another team member

# Correlation analysis (numeric columns only)
numeric_columns <- dataset %>% select_if(is.numeric)
correlation_matrix <- cor(numeric_columns, use = "complete.obs")
corrplot(correlation_matrix, method = "circle")

# Boxplot for outlier detection (Profit)
ggplot(dataset, aes(y = Profit)) +
  geom_boxplot(fill = "orange", color = "black") +
  labs(title = "Boxplot of Profit", y = "Profit")

# Pairwise scatter plots
pairs(numeric_columns)

# Loop through numeric columns to plot histograms
for(col in colnames(numeric_columns)) {
  p <- ggplot(dataset, aes_string(x = col)) +
    geom_histogram(binwidth = 10, color = "black", fill = "blue") +
    labs(title = paste("Distribution of", col), x = col, y = "Count")
  print(p)
}

#SCRIPT 2


# Coffee Chain Sales Data Analysis - Data Transformations
# This script performs a series of transformations on the "Profit" variable 
# in the Coffee Chain Sales dataset to normalize its distribution and reduce skewness. 
# Transformations applied include Log, Square Root, Box-Cox, and Yeo-Johnson. 
# The code also calculates skewness before and after transformations, allowing for 
# comparison of distribution improvements.


# Install necessary packages if not already installed
if (!require(MASS)) install.packages("MASS")
if (!require(e1071)) install.packages("e1071")
if (!require(caret)) install.packages("caret")

# Load necessary libraries
library(MASS)
library(e1071)
library(caret)

# Checking the skewness of the "Profit" variable
original_skewness <- skewness(dataset$Profit, na.rm = TRUE)
print(paste("Original Skewness of Profit:", original_skewness))


# Replace negative or NA values in Profit with a small positive value
dataset$Profit[is.na(dataset$Profit) | dataset$Profit <= 0] <- 1e-5

# Now apply the Log Transformation (adding 1 to avoid log(0) issues)
dataset$Log_Profit <- log(dataset$Profit + 1)

# Square Root Transformation
dataset$Sqrt_Profit <- sqrt(dataset$Profit)

# Box-Cox Transformation (Profit is now adjusted to be positive)
boxcox_result <- boxcox(lm(Profit ~ 1, data = dataset), plotit = FALSE)
lambda_boxcox <- boxcox_result$x[which.max(boxcox_result$y)]
dataset$BoxCox_Profit <- ((dataset$Profit ^ lambda_boxcox) - 1) / lambda_boxcox

# Yeo-Johnson Transformation (can handle zero/negative values directly)
pre_process_model <- preProcess(as.data.frame(dataset$Profit), method = "YeoJohnson")
dataset$YeoJohnson_Profit <- predict(pre_process_model, as.data.frame(dataset$Profit))

# Extract the numeric values from the result (it might return a data frame)
dataset$YeoJohnson_Profit <- as.numeric(dataset$YeoJohnson_Profit[, 1])



# Plotting histograms to visualize the transformations
par(mfrow = c(2, 2)) # Arrange plots in a 2x2 layout
hist(dataset$Log_Profit, main = "Log Transformed Profit", col = "blue", breaks = 30)
hist(dataset$Sqrt_Profit, main = "Square Root Transformed Profit", col = "green", breaks = 30)
hist(dataset$BoxCox_Profit, main = "Box-Cox Transformed Profit", col = "red", breaks = 30)
hist(dataset$YeoJohnson_Profit, main = "Yeo-Johnson Transformed Profit", col = "purple", breaks = 30)

# Calculating skewness for the original and transformed variables
log_skewness <- skewness(dataset$Log_Profit)
sqrt_skewness <- skewness(dataset$Sqrt_Profit)
boxcox_skewness <- skewness(dataset$BoxCox_Profit)
yeojohnson_skewness <- skewness(dataset$YeoJohnson_Profit)

# Display skewness of original and transformed variables
transformed_skewness <- data.frame(
  Transformation = c("Original", "Log", "Square Root", "Box-Cox", "Yeo-Johnson"),
  Skewness = c(original_skewness, log_skewness, sqrt_skewness, boxcox_skewness, yeojohnson_skewness)
)
print(transformed_skewness)
