---
title: "project"
author: "Ankita Mishra"
date: "2024-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# libraries import
library(ggplot2)
library(GGally)
library(corrplot)
library(readr)
library(leaps)
library(car)
library(MASS)
library(psych)
library(dplyr)
library(glmnet)
library(pracma)
```
```{r}
coffee = read.csv('Coffee_Chain_Sales .csv')
head(coffee)
names(coffee)[sapply(coffee, is.integer)]
names(coffee)[sapply(coffee, is.character)]
```
```{r}
# check for missing data
sum(is.na(coffee))
```
```{r}
summary(coffee)
```
```{r}
unique(coffee$Market_size)
unique(coffee$Market)
unique(coffee$Product_line)
unique(coffee$Product_type)
unique(coffee$Product)
unique(coffee$State)
unique(coffee$Type)
```


# Coffee Chain Sales Data Analysis - Exploratory Data Analysis (EDA)
# This script provides an initial exploration of the Coffee Chain Sales dataset, 
# focusing on summary statistics, missing data, correlations, outlier detection, 
# and the distributions of numeric variables. These steps help understand the data's 
# characteristics, identify potential issues, and guide subsequent analysis and 
# modeling efforts.

install.packages("ggplot2")
install.packages("dplyr")
install.packages("corrplot")

library(ggplot2)
library(dplyr)
library(corrplot)

# Loading the dataset
dataset <- read.csv('Coffee_Chain_Sales.csv')

# Summary 
summary(dataset)

# Check for missing values
missing_data <- colSums(is.na(dataset))
print(missing_data)

#missing values found : 0, fixed by another team member

# Correlation analysis (numeric columns only)
numeric_columns <- dataset %>% select_if(is.numeric)
correlation_matrix <- cor(numeric_columns, use = "complete.obs")
corrplot(correlation_matrix, method = "circle")

# Boxplot for outlier detection (Profit)
ggplot(dataset, aes(y = Profit)) +
  geom_boxplot(fill = "orange", color = "black") +
  labs(title = "Boxplot of Profit", y = "Profit")

# Pairwise scatter plots
pairs(numeric_columns)

# Loop through numeric columns to plot histograms
for(col in colnames(numeric_columns)) {
  p <- ggplot(dataset, aes_string(x = col)) +
    geom_histogram(binwidth = 10, color = "black", fill = "blue") +
    labs(title = paste("Distribution of", col), x = col, y = "Count")
  print(p)
}



